---
title: "Single-catch SCR with Times"
author: "David Borchers"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary

Simulations to check the veracity of the new MLE Ben and I came up with.

# The negative log-likelihood

```{r llik}
negll = function(pars,adists,meshdists,capthist,surveyT) {
  nocc = dim(capthist)[3]
  ntraps = dim(capthist)[2]
  n = dim(capthist)[1]
  for(j in 1:nocc) {
    
  }
}
```

# Simulations

Let $\mathbf{s}=(x,y)$ be a generic activity centre location, and $\lambda(d) = \lambda_0\frac{-d^2}{2\sigma^2} = \lambda_0\frac{-(x^2+y^2)}{2\sigma^2}$ be the hazard function evaluated at distance $d=\sqrt{x^2+y^2}$ from the activity centre, where $\lambda_0$ and $\sigma$ are the hazard function parameters.

We have $N$ activity centres, in a region of area $A$ that includes the trap array at its centre.

We assume that the times to detection of animal $i$ at trap $k$ is an exponential random variable with expectation $\lambda(d_{ik})^{-1}$, where $d_{ik}$ is the distanse from $i$'s activity centre (AC) to trap $k$.


```{r simSCR.fixedN, echo=FALSE, message=FALSE, warning=FALSE}
# Function to simulate fist detection times (and animals) at each trap:
simSCR.fixedN.single = function(locs, sigma = 0.5, lambda0 = 0.5, noccasions=1, traps, surveyT)
{
  if(dim(locs)[1]==0) stop("N==0")
  if(dim(locs)[2]!=2) stop("locs is of wrong dimension")
  N = dim(locs)[1]
  K <- nrow(traps)
  capthists <- list()
  ids <- NULL
  for(occ in 1:noccasions) {
    capthist = NULL
    for(i in 1:N) {
      d2 <- (locs[i,1] - traps[,1])^2 + (locs[i,2] - traps[,2])^2
      haz <- lambda0*exp(-d2/(2*sigma^2))
      times <- rexp(K, haz)
      capthist <- rbind(capthist, times)
    }
    capthists[[occ]] = make.single(capthist,surveyT)
  }
  
  chist = compress.capthist.list(capthists,K) # turn into n x K x noccasions array with capture times
  
  if(!is.null(chist)) {
    return(chist)
  }else {
    return(NULL)
  }
 
#  if(!is.null(capthists)) {
#    return(capthists)
#  }else {
#    return(NULL)
#  }

}

make.single = function(carray,surveyT) {
# Takes array 'carray'of capture times of all animals (rows) on all traps (cols) and turns it into
# an array with columns 'animal', 'trap' and 'time' consistent with traps being single-catch traps
  chist = NULL
  nextime = 0
  tarray = carray
  while(nextime<surveyT) {
    ind = which(tarray==min(tarray),arr.ind = TRUE) # next trap and animal
    captime = tarray[ind[1],ind[2]] # capture time
    if(captime<surveyT) {
      chist = rbind(chist,c(ind[1],ind[2],captime)) # put new capthist with existing ones
      tarray[ind[1],] = surveyT + 1 # remove this animal by making all capture times larger than survey time
      tarray[,ind[2]] = surveyT + 1 # remove this trap by making all capture times larger than survey time
    }else {
      nextime = Inf
    }
  }
  colnames(chist) = c("animal","trap","time")
  return(chist)
}

compress.capthist.list = function(capthist.list,ntraps) {
# converts a list of capture times of animals on each occasion, into a 3D array with dimensions n x K x noccasions,
# with zeros when animal i was not detected by trap k on occasion j.
  J = length(capthist.list)
  animals = NULL
  for(j in 1:J) animals = c(animals,capthist.list[[j]][,1]) # concatenate animals across occasions
  animals = sort(unique(animals)) # get unique animals across occasions
  n = length(animals) # number of animals
  # set up 3D array for capture times on all occasions:
  chist = array(rep(0,n*ntraps*J), dim=c(n,ntraps,J), dimnames=list(animal=1:n,trap=1:ntraps,occasion=1:J))
  for(j in 1:J) { # loop through occasions
    for(i in 1:n) { # loop through animals within occasions
      row = which(capthist.list[[j]][,1]==animals[i]) # find row animal i appears in (or NULL if not there)
      if(!is.null(row)) trap = capthist.list[[j]][row,2] # find trap animal was caught in, if there
      chist[i,trap,j] = capthist.list[[j]][row,3] # record capture time
    }
  }
  row.names(chist) = animals # put original animal number in as row name
  return(chist)
}
```

Set up simulation sceanrio and parameters:

```{r simsetup, message=FALSE}
library(secr)

sigma <- 0.5
s2 = sigma^2
l0 = lambda <- 6

traps <- expand.grid(x = 1:5,y = 1:5)
buffer = 3*sigma
xlim <- range(traps[,1]) + c(-buffer, buffer)
ylim <- range(traps[,2]) + c(-buffer, buffer)
a = 0.25^2
mask <- expand.grid(x = seq(xlim[1], xlim[2], sqrt(a)), y = seq(ylim[1], ylim[2], sqrt(a))  )
nmask <- nrow(mask)
J <- nrow(traps)
# Turn traps and mask into secr objects because the function I have for covariance 
# assumes that they are. Should no doubt change this in due course ...
simtraps = read.traps(data=traps, type="proximity")

trapdists = edist(simtraps,simtraps) # distance between traps

area <- nrow(mask) * a
targetD = 0.5
D = round(targetD*area)/area # to get integer N
N = D*area

surveyT = 1 # duration of each survey occasion

noccasions = 4 # number of survey occasion 

pars = log(c(D=D, lambda0=lambda, sigmasq=sigma^2))
```

So the key parameter values are as follows:

* $\sigma = `r sigma`$
* $\lambda0 = `r l0`$
* $A = `r area`$
* $N = `r N`$
* buffer = `r buffer`

Let's take a look at the traps:
```{r traplot}
ntraps = dim(traps)[1]
trapno = 1:ntraps
plot(simtraps$x,simtraps$y,col=0,
     xlim=c(min(simtraps$x)-buffer,max(simtraps$x)+buffer),
     ylim=c(min(simtraps$y)-buffer,max(simtraps$y)+buffer),
     xlab="",ylab="",main="Traps",cex.main=0.75)
text(simtraps$x,simtraps$y,labels=trapno,cex=0.5)
```

Now try simulate a survey in which N is fixed and initial locations are random but then remain fixed for repeat survey occasions. 

Simulate surveys and keep counts in a list of length `r noccasions`, with each element being an array of dimension (`r N` x `r ntraps`).
```{r dosims, cache=FALSE}
set.seed(123)
# Do survey
  # Simulate animal locations once:
  locs <- cbind(x = runif(N, xlim[1], xlim[2]), 
                y = runif(N, ylim[1], ylim[2])) # animal locations
  # Simulate noccasions sets of capture times from these locations
  capthist = simSCR.fixedN.single(locs, sigma, lambda, noccasions=noccasions, traps, surveyT=1)
```

Plot animals and traps, showing which were detected
```{r plotpopcap, fig.dim=c(8,9)}
caught.by.occasion = apply(capthist,c(1,3),sum)>0 # logical array indicating which animals (row) 
                                                  # were caught on each occasion (column)
caught.animals = as.integer(row.names(capthist)) # population number of detected animals
par(mfrow=c(2,2))
for(occ in 1:noccasions) {
  trapind = which(capthist[caught.by.occasion[,occ],,occ]>0,arr.ind = TRUE)
  trapind = trapind[order(trapind[,1]),]
  plot(simtraps$x,simtraps$y,col=0,
       xlim=c(min(simtraps$x)-buffer,max(simtraps$x)+buffer),
       ylim=c(min(simtraps$y)-buffer,max(simtraps$y)+buffer),
     xlab="",ylab="",main=paste("Occasion",occ),cex.main=0.75)
  text(simtraps$x,simtraps$y,labels=trapno,cex=0.5)
  text(locs[,1],locs[,2],labels=1:N,cex=0.5,col="red") # plot all animals
  points(locs[caught.animals,1][caught.by.occasion[,occ]], # plot animals caught on the occasion
         locs[caught.animals,2][caught.by.occasion[,occ]],col="red",cex=1.5)
  text(simtraps$x,simtraps$y,labels=trapno,cex=0.5)
  segments(locs[caught.animals,1][caught.by.occasion[,occ]],
           locs[caught.animals,2][caught.by.occasion[,occ]],
           simtraps$x[trapind[,2]], simtraps$y[trapind[,2]],
           col="gray")
}
```

Here are the first 10 rows of the capture history array:
```{r chist}
capthist[1:10,,]
```

